---
title: "Adult’s PSTRE Level Prediction"
author: "Ying Zhou"
date: "2022-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options (warn = -1)
```

##  Explore the PIAAC Dataset with R

```{r}
PIAAC<- read.csv("PIAAC_US_2012_subsample.csv", header=TRUE)

library(dplyr)

PIAAC1<- select(PIAAC,+contains("num_of_diff_emails_views"),+contains("num_of_email_views"),+contains("num_switching_environment"),+contains("num_of_diff_visited_pages"),+contains("num_of_page_visit"),+contains("time_on_task"),"PVLIT1","PVNUM1")

PIAAC1["U"]<-PIAAC["U01a000A"]+PIAAC["U01b000A"]+PIAAC["U03a000A"]+PIAAC["U06a000A"]+PIAAC["U06b000A"]+PIAAC["U21x000A"]+PIAAC["U04a000A"]+PIAAC["U19a000A"]+PIAAC["U19b000A"]+PIAAC["U07x000A"]+PIAAC["U02x000A"]+PIAAC["U16x000A"]+PIAAC["U11b000A"]+PIAAC["U23x000A"]

#PIAAC1<-scale(PIAAC1,center=TRUE, scale=TRUE)*10
PS_class<-PIAAC["PS_class"]
PIAAC1<-cbind(PIAAC1,new_col=PS_class)
PIAAC1<-na.omit(PIAAC1)
#names(PIAAC1)
#skim(PIAAC)
```

## Setting up training and test set
```{r}
PIAAC1$PS_class<-as.factor(PIAAC1$PS_class)
library(caret)

set.seed(9)

inTrain <- createDataPartition(y = PIAAC1$PS_class, p = .8, list = FALSE)
PIAAC1_train <- PIAAC1 %>% slice(inTrain)
PIAAC1_test <- PIAAC1 %>% slice(-inTrain)

train_index <- createFolds(PIAAC1_train$PS_class, k = 10) ##10 fold inner loop cross-validation
```


## Random Forest

```{r}
randomForestFit <- PIAAC1_train %>% train(PS_class ~ .,
  method = "rf",
  data = .,
    tuneLength = 5,
    trControl = trainControl(method = "cv", indexOut = train_index))
randomForestFit
```

```{r}
plot(randomForestFit$finalModel, main="Random Forest Model")
```


```{r}
print(varImp(randomForestFit))
plot(varImp(randomForestFit))
```



```{r}
pr_rf <- predict(randomForestFit, PIAAC1_test)
pr_rf
```


```{r}
#confusionMatrix(pr_rf, reference = PIAAC1_test$PS_class)
confusionMatrix(pr_rf, reference = PIAAC1_test$PS_class, mode = "prec_recall")
```

## Linear Support Vector Machine

```{r}
svmFit <- PIAAC1_train %>% train(PS_class ~.,
  method = "svmLinear",
  data = .,
    tuneLength = 5,
    trControl = trainControl(method = "cv", indexOut = train_index))
svmFit
```


```{r}
print(varImp(svmFit))
#plot(varImp(svmFit))
```


Make the prediction on test data

```{r}
pr_svm <- predict(svmFit, PIAAC1_test)
pr_svm
```

Check confusion matrix between prediction and actual type (gold standard) in the test set.

```{r}
#confusionMatrix(pr_svm, reference = PIAAC1_test$PS_class)
confusionMatrix(pr_svm, reference = PIAAC1_test$PS_class, mode = "prec_recall")  ##get the precision, recall and F-measure
```


## Support Vector Machine with Radial Kernel

```{r}
svmrad <- PIAAC1_train %>% train(PS_class ~.,
  method = "svmRadial",
  data = .,
    tuneLength = 5,
    trControl = trainControl(method = "cv", indexOut = train_index))
svmrad
```


```{r}
print(varImp(svmrad))
#plot(varImp(svmrad))
```


```{r}
pr_svmrad <- predict(svmrad, PIAAC1_test)
pr_svmrad
```


```{r}
confusionMatrix(pr_svmrad, reference = PIAAC1_test$PS_class)
```


## Conditional Inference Tree (Decision Tree)
```{r}
ctreeFit <- PIAAC1_train %>% train(PS_class ~ .,
  method = "ctree",
  data = .,
    tuneLength = 5,
    trControl = trainControl(method = "cv", indexOut = train_index))
ctreeFit
```


```{r}
print(varImp(ctreeFit))
#plot(varImp(ctreeFit))
```


Make the prediction on test data

```{r}
pr_ctree <- predict(ctreeFit, PIAAC1_test)
pr_ctree
```

Check confusion matrix between prediction and actual type (gold standard) in the test set.
```{r}
confusionMatrix(pr_ctree, reference = PIAAC1_test$PS_class)
```




